---
title: "CLEAN Logistic + Nonlinear Analysis [Updated April 2020]"
author: "Will Simmons"
date: "04/06/2020"
output: 
  html_document:
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

<!-- Setup -->

```{r}

shh = suppressMessages

shh(library(tidyverse))
shh(library(cowplot))
shh(library(patchwork))
shh(library(viridis))
shh(library(mgcv))
shh(library(conflicted))
conflict_prefer("filter", "dplyr")

new_data = readRDS('./data/new_analytic_dataset.RDS')

knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
options(width = 1000)

```

<!-- Models -->

<!-- Exposure lists -->
```{r}

# Exposure list
log_exposure_list_creatinine =
  new_data %>% 
  select(starts_with("log_") & ends_with("_creatinine")) %>% 
  names() %>% 
  as_tibble() %>% 
  rename(predictor_string = value) %>% 
  mutate(predictor_string = as.character(predictor_string))

# Exposures as character vector
exposures =
  log_exposure_list_creatinine %>% 
  as.list() %>% unlist()

# Covariate/confounder set
covariate_list =
  c("sex",
    "age",
    "race",
    "hh_income"
  )

```

<!-- Crude linear models -->
```{r}

uni_ols = function(predictor) {
  
   lm(reformulate(as.character(predictor),
                  'log(tsh)'),
      data = new_data)
    
}

crude_ols_models = 
  map(exposures, uni_ols)

```


<!-- Adjusted linear models -->
```{r}

adj_ols = function(predictor) {
  
  covariates = covariate_list
  
  lm(reformulate(c(as.character(predictor), covariates),
                  'log(tsh)'),
     data = new_data) 
  
}

# Example
# adj_ols('bpa_creatinine')

# Map over all
adj_ols_models =
  map(exposures, adj_ols)

```

<!-- Adjusted linear models, created with gam() to compare with nonlinear models via GCV score -->
```{r}

adj_ols_gam = function(predictor) {
  
  covariates = covariate_list
  
  gam(reformulate(c(as.character(predictor), covariates),
                  'log(tsh)'),
      data = new_data) 
  
}

# Example
# adj_ols('bpa_creatinine')

# Map over all
adj_ols_models_gam =
  map(exposures, adj_ols_gam)

```

# Justification of log-log - (1) example diagnostic plot

After evaluating model residuals (full diagnostic analyses not shown), we decided to log-transform both our exposure chemicals and our outcome, TSH (log-log transformation). We found that this resolved issues of heteroskedasticity and provided reasonable interpretability. For an example of diagnostic plots before and after log-log transformation, see below.

<!-- Diagnostic plots -->
```{r}

# Without log-log: example
# see ./figures/non_logged_qq.png

# With log-log: example
plot(adj_ols_models[[1]], which = 2)
title("Exposure: log creatinine-adjusted 4-tert-octylphenol; Outcome: log serum TSH")
# also see ./figures/log_qq.png

```

# Correlation matrix

<!-- Correlation matrix -->
```{r}

new_data %>% #[exposures] %>% view()
  select(log_tsh, unname(exposures), sex, age, hh_income) %>% 
  mutate(hh_income = as.numeric(hh_income),
         sex = as.numeric(sex)) %>% 
  cor() %>% 
  ggcorrplot::ggcorrplot(type = "lower",
                         colors = c("#ED6A5A", "#FFFFF9", "#36C9C6"),
                         show.diag = FALSE,
                         lab = TRUE,
                         lab_size = 4) +
                         labs(title = "Figure 1. Correlation matrix of outcome and features") +
                         theme(plot.title.position = "plot",
                               legend.position = "bottom")

```


# Crude log-log OLS (8) - table of results

First, we fit eight unadjusted linear regression models, one for each of our eight exposure chemicals of interest. The outcome, TSH, remained constant across models. Both exposure and outcome were log-transformed. All adjusted models included sex, age, race, and household income to control for potential confounding.

```{r}
library(knitr)
library(kableExtra)

crude_ols_tidy =
  map_dfr(crude_ols_models, broom::tidy)

crude_ols_table =
  crude_ols_tidy %>% 
  rename("Outcome: log serum TSH" = term) %>% 
  kable(caption = "Crude linear models") %>% 
  kable_styling(latex_options = "striped") %>% 
  pack_rows(
    index = c(
      "Model 1" = 2,
      "Model 2" = 2,
      "Model 3" = 2,
      "Model 4" = 2,
      "Model 5" = 2,
      "Model 6" = 2,
      "Model 7" = 2,
      "Model 8" = 2
    )
  ) 
  
# crude_ols_table %>% save_kable('./figures/crude_table.png')

```

[table of crude results]

# Adjusted log-log OLS (8) - table of results

```{r}

adj_ols_tidy =
  map_dfr(adj_ols_models, broom::tidy)

adj_ols_table =
  adj_ols_tidy %>% 
  filter(str_detect(term, pattern = "Int") | 
           str_detect(term, pattern = "log")) %>% 
  rename("Outcome: log serum TSH" = term) %>% 
  kable(caption = "Adjusted linear models") %>% 
  kable_styling(latex_options = "striped") %>% 
  pack_rows(
    index = c(
      "Model 1" = 2,
      "Model 2" = 2,
      "Model 3" = 2,
      "Model 4" = 2,
      "Model 5" = 2,
      "Model 6" = 2,
      "Model 7" = 2,
      "Model 8" = 2
    )
  ) %>% 
  footnote(
    general_title = " ",
    "All models adjusted for sex, age, race, and household income."
  )

# adj_ols_table %>% save_kable('./figures/adj_table.png')

```

# Adjusted nonlinear (8) - table of results

```{r}

adj_nonlinear = function(predictor) {
  
  covariates = covariate_list
  
  gam(reformulate(c(paste0("s(", predictor, ")"), covariates),
                  "log(tsh)"),
      data = new_data)
  
}

# Example
# adj_nonlinear("triclosan_creatinine")

adj_nonlinear_models = 
  map(exposures, adj_nonlinear)

adj_nonlinear_tidy =
  map_dfr(adj_nonlinear_models, broom::tidy)

adj_nonlinear_table =
  adj_nonlinear_tidy %>% 
  filter(str_detect(term, pattern = "Int") | 
           str_detect(term, pattern = "log")) %>% 
  rename("Outcome: log serum TSH" = term) %>% 
  kable(caption = "Adjusted general additive models with penalized thin-plate spline terms applied to log chemical exposures<br>
        Table of nonparametric spline results",
        escape = F) %>% 
  kable_styling() %>% 
  pack_rows(
    index = c(
      "Model 1" = 1,
      "Model 2" = 1,
      "Model 3" = 1,
      "Model 4" = 1,
      "Model 5" = 1,
      "Model 6" = 1,
      "Model 7" = 1,
      "Model 8" = 1
    )
  ) %>% 
  footnote(
    general_title = " ",
    "All models adjusted for sex, age, race, and household income."
  )

# adj_nonlinear_table %>% save_kable('./figures/nonlinear_table.png')

```


# Plots:

  ## If all OLS, forest plots of effect estimates for eight adjusted models
  
  ## If any nonlinearities, plots of nonlinear function (predicted values?)

<!-- Plot of nonlinearities for benzophenone-3 and methyl paraben   -->
```{r}

plot(adj_nonlinear_models[[2]])
title("Adjusted penalized spline: log benzophenone-3 (Outcome: log serum TSH)")

plot(adj_nonlinear_models[[7]])
title("Adjusted penalized spline: methyl paraben (Outcome: log serum TSH)")

```
  
# Comparing models with GCV score

```{r}

# Models 2 and 7

adj_ols_models_gam[[2]]$gcv.ubre
adj_nonlinear_models[[2]]$gcv.ubre

adj_ols_models_gam[[7]]$gcv.ubre
adj_nonlinear_models[[7]]$gcv.ubre

```

# Trying natural spline term in lm()

```{r}

library(splines)

# Compare to adj_ols_models[[2]]
adj_nat_spline_2_2df = lm(
  data = new_data,
  formula = log(tsh) ~ ns(log_benzophenone_3_creatinine, df = 2) + sex + age + race + hh_income
)
AIC(adj_nat_spline_2_2df)
# AIC = 3576.504

adj_nat_spline_2_3df = lm(
  data = new_data,
  formula = log(tsh) ~ ns(log_benzophenone_3_creatinine, df = 3) + sex + age + race + hh_income
)
AIC(adj_nat_spline_2_3df)
# AIC = 3573.706*** 

adj_nat_spline_2_4df = lm(
  data = new_data,
  formula = log(tsh) ~ ns(log_benzophenone_3_creatinine, df = 4) + sex + age + race + hh_income
)
AIC(adj_nat_spline_2_4df)
# AIC = 3574.914

# Compare to adj_ols_models[[7]]
adj_nat_spline_7_2df = lm(
  data = new_data,
  formula = log(tsh) ~ ns(log_methyl_paraben_creatinine, df = 2) + sex + age + race + hh_income
)
AIC(adj_nat_spline_7_2df)
# AIC = 3577.454

adj_nat_spline_7_3df = lm(
  data = new_data,
  formula = log(tsh) ~ ns(log_methyl_paraben_creatinine, df = 3) + sex + age + race + hh_income
)
AIC(adj_nat_spline_7_3df)
# AIC = 3575.892***

adj_nat_spline_7_4df = lm(
  data = new_data,
  formula = log(tsh) ~ ns(log_methyl_paraben_creatinine, df = 4) + sex + age + race + hh_income
)
AIC(adj_nat_spline_7_4df)
# AIC = 3576.337

```

# Predicting and plotting both df=3 natural splines

```{r}

# Model 2
predictions_2 =
  predict(adj_nat_spline_2_3df,
          se.fit = TRUE,
          type = 'terms') %>% 
  as.data.frame() %>% 
  as_tibble() %>% 
  select(
    fit = fit.ns.log_benzophenone_3_creatinine..df...3.,
    se = se.fit.ns.log_benzophenone_3_creatinine..df...3.
  ) %>% 
  mutate(
    lb = fit - se*1.96,
    ub = fit + se*1.96,
    x = new_data$log_benzophenone_3_creatinine
  )

ggplot(data = predictions_2, aes(x = x, y = fit)) +
  geom_line() +
  geom_line(aes(y = lb)) +
  geom_line(aes(y = ub)) +
  geom_rug(sides = "b")


# Model 7
predictions_7 = 
  predict(adj_nat_spline_7_3df,
          se.fit = TRUE,
          type = 'terms') %>% 
  as.data.frame() %>% 
  as_tibble() %>% 
  select(
    fit = fit.ns.log_methyl_paraben_creatinine..df...3.,
    se = se.fit.ns.log_methyl_paraben_creatinine..df...3.
  ) %>% 
  mutate(
    lb = fit - se*1.96,
    ub = fit + se*1.96,
    x = new_data$log_methyl_paraben_creatinine
  )

ggplot(data = predictions_7, aes(x = x, y = fit)) +
  geom_line() +
  geom_line(aes(y = lb)) +
  geom_line(aes(y = ub))

```

# Plotting predicted values with linear regression predictions

```{r}

predictions_ols_2 =
  predict(adj_ols_models[[2]],
          se.fit = TRUE,
          type = 'terms') %>% 
  as.data.frame() %>% 
  as_tibble() %>% 
  select(
    fit = fit.log_benzophenone_3_creatinine,
    se = se.fit.log_benzophenone_3_creatinine
  ) %>% 
  mutate(
    lb = fit - se*1.96,
    ub = fit + se*1.96,
    x = new_data$log_benzophenone_3_creatinine
  )

ggplot(data = predictions_ols_2, aes(x = x, y = fit)) +
  geom_line() +
  geom_line(aes(y = lb)) +
  geom_line(aes(y = ub)) +
  
  geom_line(data = predictions_2, aes(x = x, y = fit), color = 'indianred') +
  geom_line(data = predictions_2, aes(y = lb), color = 'indianred', alpha = 0.4) +
  geom_line(data = predictions_2, aes(y = ub), color = 'indianred', alpha = 0.4)

predictions_ols_7 =
  predict(adj_ols_models[[7]],
          se.fit = TRUE,
          type = 'terms') %>% 
  as.data.frame() %>% 
  as_tibble() %>% 
  select(
    fit = fit.log_methyl_paraben_creatinine,
    se = se.fit.log_methyl_paraben_creatinine
  ) %>% 
  mutate(
    lb = fit - se*1.96,
    ub = fit + se*1.96,
    x = new_data$log_methyl_paraben_creatinine
  )

ggplot(data = predictions_ols_7, aes(x = x, y = fit)) +
  geom_line() +
  geom_line(aes(y = lb)) +
  geom_line(aes(y = ub)) +
  
  geom_line(data = predictions_7, aes(x = x, y = fit), color = 'indianred') +
  geom_line(data = predictions_7, aes(y = lb), color = 'indianred', alpha = 0.4) +
  geom_line(data = predictions_7, aes(y = ub), color = 'indianred', alpha = 0.4)

```

# Predicting GAM for final nonlinear figure

```{r}

# GAM 2
predictions_gam_2 =
  predict(adj_nonlinear_models[[2]],
          se.fit = TRUE,
          type = 'terms') %>% 
  as.data.frame() %>% 
  as_tibble() %>% 
  select(
    fit = fit.s.log_benzophenone_3_creatinine.,
    se = se.fit.s.log_benzophenone_3_creatinine.
  ) %>% 
  mutate(
    lb = fit - se*1.96,
    ub = fit + se*1.96,
    x = new_data$log_benzophenone_3_creatinine
  )

ggplot(data = predictions_ols_2, aes(x = x, y = fit)) +
  geom_line(color = 'indianred', size = .7) +
  geom_line(aes(y = lb), color = 'gray') + 
  geom_line(aes(y = ub), color = 'gray') + 
  
  geom_line(data = predictions_gam_2, aes(x = x, y = fit), size = 1) +
  geom_line(data = predictions_gam_2, aes(y = lb), size = .8) +
  geom_line(data = predictions_gam_2, aes(y = ub), size = .8) +
  
  theme_bw()


# GAM 7
predictions_gam_7 =
  predict(adj_nonlinear_models[[7]],
          se.fit = TRUE,
          type = 'terms') %>% 
  as.data.frame() %>% 
  as_tibble() %>% 
  select(
    fit = fit.s.log_methyl_paraben_creatinine.,
    se = se.fit.s.log_methyl_paraben_creatinine.
  ) %>% 
  mutate(
    lb = fit - se*1.96,
    ub = fit + se*1.96,
    x = new_data$log_methyl_paraben_creatinine
  )

ggplot(data = predictions_ols_7, aes(x = x, y = fit), size = 1) +
  geom_line(color = 'indianred', size = .7) +
  geom_line(aes(y = lb), color = 'gray') +
  geom_line(aes(y = ub), color = 'gray') +
  
  geom_line(data = predictions_gam_7, aes(x = x, y = fit)) +
  geom_line(data = predictions_gam_7, aes(y = lb), size = .8) +
  geom_line(data = predictions_gam_7, aes(y = ub), size = .8) +
  
  theme_bw()

```

# Final Figure 2

```{r}

plot_2 =
  ggplot(data = predictions_ols_2, aes(x = x, y = fit)) +
  geom_line(color = 'indianred', size = .7) +
  geom_line(aes(y = lb), color = 'indianred', size = .4, linetype = 'dotted') +
  geom_line(aes(y = ub), color = 'indianred', size = .4, linetype = 'dotted') +
  
  geom_line(data = predictions_gam_2, aes(x = x, y = fit), size = 1) +
  geom_line(data = predictions_gam_2, aes(y = lb), size = .8) +
  geom_line(data = predictions_gam_2, aes(y = ub), size = .8) +
  
  theme_bw() +
  ggtitle("A.") +
  labs(x = "log benzophenone-3 (μg/g creatinine)",
       y = "log thyroid stimulating hormone (uIU/mL)") +
  ylim(-0.5, 0.6)


plot_7 =
  ggplot(data = predictions_ols_7, aes(x = x, y = fit)) +
  geom_line(color = 'indianred', size = .7) +
  geom_line(aes(y = lb), color = 'indianred', size = .4, linetype = 'dotted') +
  geom_line(aes(y = ub), color = 'indianred', size = .4, linetype = 'dotted') +
  
  geom_line(data = predictions_gam_7, aes(x = x, y = fit), size = 1) +
  geom_line(data = predictions_gam_7, aes(y = lb), size = .8) +
  geom_line(data = predictions_gam_7, aes(y = ub), size = .8) +
  
  theme_bw() +
  ggtitle("B.") +
  labs(x = "log methyl paraben (μg/g creatinine)",
       y = "") +
  ylim(-0.5, 0.6) 

figure_3 =
  cowplot::plot_grid(plotlist = list(plot_2, plot_7)) %>%
  patchwork::wrap_elements() +
  ggtitle(
    "Figure 3. Linear (red) and penalized spline (black) terms for log exposure values",
    subtitle = "Plotted against log thyroid stimulating hormone, with 95% confidence intervals"
  ) +
  labs(caption = "\nLinear and penalized spline models adjusted for creatinine, sex, age, race, and household income.") +
  theme(plot.caption = element_text(hjust = 0))

library(Cairo)
ggsave(plot = figure_3,
       './figures/figure_3.png',
       dpi = 300,
       width = 10, height = 6,
       type = 'cairo')
```

