---
title: "New Data Import"
author: "Will Simmons"
date: "4/1/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

FINAL DATASET

  OUTCOME
    * tsh
    
    + Outcome is not logged

  PREDICTORS ******** still determining whether these need to be logged *******
    * tert_octylphenol_creatinine
    * benzophenone_3_creatinine
    * bpa_creatinine
    * triclosan_creatinine
    * butyl_paraben_creatinine
    * ethyl_paraben_creatinine
    * methyl_paraben_creatinine
    * propyl_paraben_creatinine
    
    + All predictors transformed as follows:
      * Divided by individual creatinine value
      * Transformed via log10?                          
  
  COVARIATES
    * bmi (?) (Marianthi: could be mediator, check on this)
    * sex (1 = male, 2 = female)
    * age (continuous)
    * race (see NHANES for categories)
    * cotinine (marker of tobacco smoke exposure)
    * hh_income (household income, 15 categories - see NHANES)
    
    + For specific categories, etc. look at NHANES site for variable name (e.g. variable name for hh_income is actually INDHHIN2 - see code below where this is renamed from the original NHANES data)
  

```{r}
library(tidyverse)
library(haven)
```


```{r}

thyroid_7 = read_xpt('./data/THYROD_07_08.XPT')
thyroid_9 = read_xpt('./data/THYROD_09_10.XPT')

demo_7 = read_xpt('./data/DEMO_07_08.XPT')
demo_9 = read_xpt('./data/DEMO_09_10.XPT')
  
eph_7 = read_xpt('./data/EPH_07_08.XPT') %>% select(-URXUCR)
eph_9 = read_xpt('./data/EPH_09_10.XPT')

bmi_7 = read_xpt('./data/BMX_07_08.XPT')

cot_7 = read_xpt('./data/COTNAL_07_08.XPT')

med_7 = read_xpt('./data/MCQ_07_08.XPT')

preg_7 = read_xpt('./data/UCPREG_07_08.XPT')

```

Creating year-specific DFs

```{r}

data_7 =
  list(demo_7, eph_7, thyroid_7, bmi_7, cot_7, med_7, preg_7) %>%
  reduce(full_join, by = "SEQN") %>% 
  mutate(year = "2007-2008")

data_9 =
  list(demo_9, eph_9, thyroid_9) %>%
  reduce(full_join, by = "SEQN") %>% 
  mutate(year = "2009-2010")

```

Selecting variables from joined data

```{r}

joined = 
  full_join(
    data_7,
    data_9
  ) %>%  
  select(
    
    # Covariates:
      ## BMI
    bmi = BMXBMI,
      ## x Household income
    hh_income = INDHHIN2,
      ## x Creatinine
      ## x Age
      ## x Sex
      ## x Race/ethnicity
      ## Iodine will NOT be covariate - see Rousset, 1981 (potential mediator for methyl paraben --> thyroid function)
      ## Cotinine? Measure of exposure to tobacco smoke - potentially confounder
    cotinine = LBXCOT,
    cotinine_det = LBDCOTLC,
    creatinine = URXUCR,
    
    # Need exclusionary variables:
      ## Pregnant 
    preg = URXPREG, # if == 1, exclude
    
      ## History of thyroid disease
    thy_history = MCQ160M,   # if == 1, exclude
    
    # General
    id = SEQN,
    year,
    
    # From thyroid
    thyroglobulin_antibodies = LBXATG,
    t3_free = LBXT3F,
    thyroxine_free = LBXT4F,
    thyroglobulin = LBXTGN,
    tsh = LBXTSH1,
    peroxidase_antibodies = LBXTPO,
    t3_total = LBXTT3,
    thyroxine_total = LBXTT4,
    
    # From EPH
    tert_octylphenol = URX4TO,
    tert_octylphenol_det = URD4TOLC,
    benzophenone_3 = URXBP3,
    benzophenone_3_det = URDBP3LC,
    bpa = URXBPH,
    bpa_det = URDBPHLC,
    triclosan = URXTRS,
    triclosan_det = URDTRSLC,
    butyl_paraben = URXBUP,  
    butyl_paraben_det = URDBUPLC, 
    ethyl_paraben = URXEPB,  
    ethyl_paraben_det = URDEPBLC, 
    methyl_paraben = URXMPB,
    methyl_paraben_det = URDMPBLC,
    propyl_paraben = URXPPB,
    propyl_paraben_det = URDPPBLC,
    
    # From demographic
    sex = RIAGENDR,
    age = RIDAGEYR,
    race = RIDRETH1
    
  ) %>% 
  
  mutate(
    preg = as.factor(preg),
    thy_history = as.factor(thy_history)
  ) %>% 
  
  filter(
    # Pregnant
    (preg != 1 | is.na(preg)),
    # Thyroid history
    (thy_history != 1 | is.na(thy_history)),
    # No missing income data
    !(hh_income %in% c(77,99)),
  ) %>% 
  select(
    -preg, -thy_history
  ) %>% 
  mutate_at(
    vars(c("hh_income", "sex", "race")),
    funs(as.factor(.))
  ) %>% 
  mutate_at(
    vars(ends_with("_det")),
    funs(as.factor(.))
  ) %>% 
  mutate(id = as.character(id))

# colSums(is.na(joined))
# Amelia::missmap(joined)
# # creatinine missing from all obs - in both EPH and cotinine data (used cotinine version, fewer missing)
# eph_9 %>% select(URXUCR)
# 
# colSums(is.na(joined %>% filter(year == "2009-2010")))
# 
# joined %>% filter(year == "2009-2010")

joined_nomiss =
  joined %>% drop_na() %>% 
  # Leaves 2,018 observations
  # Ugh all from 2007-2008, since no one in 2009-2010 had both phenols and thyroid measures
  
  # Adjusting for creatinine
  mutate(
    tert_octylphenol_creatinine = tert_octylphenol/(creatinine*10),
    benzophenone_3_creatinine = benzophenone_3/(creatinine*10),
    bpa_creatinine = bpa/(creatinine*10),
    triclosan_creatinine = triclosan/(creatinine*10),
    butyl_paraben_creatinine = butyl_paraben/(creatinine*10),
    ethyl_paraben_creatinine = ethyl_paraben/(creatinine*10),
    methyl_paraben_creatinine = methyl_paraben/(creatinine*10),
    propyl_paraben_creatinine = propyl_paraben/(creatinine*10)
    
  )


```

Saving final dataset

```{r}

saveRDS(joined_nomiss, './data/new_analytic_dataset.RDS')
# data = readRDS('./data/new_analytic_dataset.RDS')

```

