---
title: "Logistic + Nonlinear Analysis"
author: "Will Simmons"
date: "3/24/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

# 0. Importing and checking LODs
```{r}

library(tidyverse)
library(conflicted)

data = readRDS('./data/analytic_dataset.RDS')

```

```{r, fig.width = 13, fig.height = 8}
library(viridis)
lod =
  data %>% 
  pivot_longer(
    matches("det$"),
    names_to = "bisphenol",
    values_to = "lod"
  ) %>% 
  select(year, bisphenol, lod) %>% 
  mutate(
    bisphenol = str_replace(bisphenol, "_u_det", "")
  ) %>% 
  mutate(
    lod = as.factor(lod),
    lod = fct_recode(lod, 
                     "Above LOD" = "0",
                     "Below LOD" = "1"),
    year = fct_recode(year, 
                      "2013-14" = "13_14",
                      "2015-16" = "15_16")
  ) %>% 
  group_by(year, bisphenol, lod) %>% 
  summarize(count = n()) %>% 
  ggplot(aes(fill = lod, x = bisphenol, y = count)) +
    geom_bar(position = "dodge", stat = "identity") +
    theme_bw() +
    scale_fill_brewer(palette = 5) +
    scale_x_discrete(labels = c("bpa" = "BPA", "bpf" = "BPF", "bps" = "BPS")) +
    facet_wrap(~year, ncol = 1) +
    theme(legend.position = "bottom",
          legend.title = element_blank(),
          plot.title = element_text(hjust = 0.5)) +
    labs(x = "",
         y = "Observations",
         title = "Distribution of Observations by LOD and Survey Period")

lod

```

It looks like the following variables likely have too high of a proportion of LOD:

  * BPF
  * Butyl paraben
  * Ethyl paraben
  * Triclocarban
  * Triclosan
  
Which leaves us with:

  * BPA
  * BPS
  * 2,4-dichlorophenol
  * 2,5-dichlorophenol
  * Methyl paraben
  * Propyl paraben
  
Let's look at how correlated our remaining exposures are:

```{r}
# Correlation of exposures
data %>% 
  select(
    bpa_u,
    bps_u,
    matches("^dichl"),
    starts_with(c("methyl", "propyl")),
    -matches("det$")
  ) %>% 
  cor() %>% 
  corrplot::corrplot(type = "lower",
                     method = "square", 
                     addCoef.col = "black", 
                     diag = FALSE, 
                     number.cex = .6,
                     tl.col = "black",
                     tl.cex = .9,
                     tl.srt = 45)

```

Methyl and propyl paraben are highly correlated (r = 0.69), as are the two dichlorophenol measures (r = 0.97). Thus, we're essentially left with BPA, BPS, a paraben (one of methyl/propyl), and a dichlorophenol (one of 2,4- and 2,5-).

# 1. Crude logistic regression models

```{r}
# Creating function to run univariable logistic regression
uni_logistic = function(predictor) {
  
  glm(reformulate(as.character(predictor),
                  'thyroid_outcome'),
      data = data,
      family = 'binomial') %>% 
  
  broom::tidy()
  
}

# Example
# uni_logistic("bpa_u")
```

```{r}
# List of covariate names to map over
all_variables =
  data %>% 
  select(
    -year,
    -id,
    -ends_with("_det"),
    -thyroid_outcome
  ) %>% 
  names() %>% 
  as.list()

# Dataframe of univariable logistic regression models
uni_results =
  map_dfr(all_variables, uni_logistic)

```

# 2. Adjusted logistic regression models

```{r}
# Covariate/confounder set
covariates =
  c(
    "sex",
    "age",
    "race",
    "year",
    "creatinine_u"
  )
  
# Exposure set
exposures =
  data %>% 
  select(
    ends_with("u"),
    -starts_with("creatinine")
  ) %>% 
  names() %>% 
  as.list() 

```


# 3. Crude nonlinear models

# 4. Adjusted nonlinear models

